---
title: "COMPSCIX 415.2 Homework 6"
author: "Shannon Hamilton"
date: "3/11/2018"
output: 
  html_document: 
    toc: yes
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(mosaicData)
```

##Exercise 1

<i>1. What variables are in this data set?</i><p></p>

<p>The variables in this data scet are (1) the survival status of each woman in the data after 20 years, (2) if each woman is a smoker or not, and (3) each woman's age at the time of the initial survey. </p>

<i>2. How many observations are there and what does each represent?</i><p></p>

<p>There are 1314 observations in the data. Each represents one woman. </p>

<i>3. Create a table (use the R code below as a guide) and a visualization of the relationship between smoking
status and outcome, ignoring age. What do you see? Does it make sense?</i><p></p>

<p>We see that there are overall more non-smokers than smokers in the data, and more of the non-smokers proportionately were dead at the end of the study than smokers, which is unexpected. I would expect more smokers to be dead, but we are not controlling for age. </p>

```{r}
data("Whickham") 

Whickham %>%
count(smoker , outcome )

ggplot(Whickham) + 
  geom_bar(mapping = aes(x = smoker, fill = outcome))

```

<i>4. Recode the age variable into an ordered factor with three categories: age <= 44, age > 44 & age <=64, and age > 64. Now, recreate visualization from above, but facet on your new age factor. What do you see? Does it make sense?</i><p></p>


```{r}
data("Whickham") 

Wickham1 <- Whickham %>%
mutate(age2 = cut(age, breaks=c(-Inf, 44, 64, Inf), labels=c("<=44","44_64",">64")))

Wickham1 %>%
count(age2, smoker , outcome )

ggplot(Wickham1) +
geom_bar(mapping = aes(x = smoker, fill = outcome)) +
facet_wrap(~ age2, nrow=2)

 

```

<p>Now the trend aligns to my expectations once I include age as a factor in the analysis.</p>


##Exercise 2


<i>1. Generate a random sample of size n = 10000 from a gamma(1,2) distribution and plot a histogram or density curve. Use the code below to help you get your sample</i><p></p>

```{r}
n <- 10000
# look at ?rgamma to read about this function
(gamma_samp <- tibble(x = rgamma(n, shape = 1, scale = 2)))


ggplot(gamma_samp, aes(x = x, y = ..count..)) +
geom_histogram()

```


<i>2. What is the mean and standard deviation of your sample? They should both be close to 2 because for
a gamma distribution:</i><p></p>

```{r}


(mean_samp <- mean(gamma_samp$x))
(sd_samp <- sd(gamma_samp$x))

```

<p> The mean is 1.99 and the standard deviation is 2.03, both of which are close to 2. </p>


<i>3. Pretend the distribution of our population of data looks like the plot above. Now take a sample of size n = 30 from a Gamma(1,2) distribution, plot the histogram or density curve, and calculate the mean and standard deviation.</i><p></p>

```{r}
n <- 30
gamma_samp2 <- tibble(x = rgamma(n, shape = 1, scale = 2))


ggplot(gamma_samp2, aes(x = x, y = ..count..)) +
geom_histogram()


(mean_samp2 <- mean(gamma_samp2$x))
(sd_samp2 <- sd(gamma_samp2$x))

```

<i>4. Take a sample of size n = 30, again from the Gamma(1,2) distribution, calculate the mean, and assign
it to a vector named mean_samp. Repeat this 10000 times!!!! The code below might help.</i><p></p>


```{r}
# create a vector with 10000 NAs
mean_samp <- rep(NA, 10000)

# start a loop
for(i in 1:10000) {
g_samp <- rgamma(30, shape = 1, scale = 2)
mean_samp[i] <- mean(g_samp)
}
# Convert vector to a tibble
(mean_samp <- tibble(mean_samp))

```


<i>5.Make a histogram of your collection of means from above (mean_samp).</i><p></p>

```{r}


ggplot(mean_samp, aes(x = mean_samp, y = ..count..)) +
geom_histogram()

```


<i>6. Calculate the mean and standard deviation of all of your sample means. </i><p></p>

```{r}

(mean_samp3 <- mean(mean_samp$mean_samp))
(sd_samp3 <- sd(mean_samp$mean_samp))

```

<i>7. Did anything surprise you about your answers to #6? </i><p></p>

<p> When you look across the means of many samples of a small size (30), then the mean of the means approaches that of what we would expect from the distribution and the same as using one large sample. </p>


<i>8. According to the Central Limit Theorem, the mean of your sampling distribution should be very close
to 2, and the standard deviation of your sampling distribution should be close to σ/√n = 2/√30 = 0.365.
Repeat #4-#6, but now with a sample of size n = 300 instead. Do your results match up well with the
theorem?</i><p></p>

```{r}

# create a vector with 10000 NAs
mean_samp <- rep(NA, 10000)

# start a loop
for(i in 1:10000) {
g_samp <- rgamma(300, shape = 1, scale = 2)
mean_samp[i] <- mean(g_samp)
}
# Convert vector to a tibble
(mean_samp <- tibble(mean_samp))


ggplot(mean_samp, aes(x = mean_samp, y = ..count..)) +
geom_histogram()

(mean_samp4 <- mean(mean_samp$mean_samp))
(sd_samp4 <- sd(mean_samp$mean_samp))
```

<p>Yes, the mean and standard deviations meet expectaions of the CLT for a sample size of 300.</p>


