---
title: "Homework 8"
author: "Shannon Hamilton"
date: "3/25/2018"
output: 
  html_document: 
    toc: yes
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
```

##Exercise 1

<i>Load the train.csv dataset into R. How many observations and columns are there? Convert the target variable
to a factor because it will be loaded into R as an integer by default.</i><p></p>

<p>There are 891 observatons and 12 variables/columns. </p>


```{r}
file_path <- '/Users/shamilto/Downloads/train-1.csv'
train <-read_csv(file_path,col_names=TRUE)
train <- as.tibble(train)
train$Survived <- as.factor(train$Survived)
glimpse(train)

```

##Exercise 2

<p><i>Our first step is to randomly split the data into train and test datasets. We will use a 70/30 split, and use
the random seed of 29283 so that we all should get the same training and test set.</p></i>


```{r}
set.seed(29283)

train_set <- train %>% sample_frac(0.7)

test_set <- train %>% filter(!(PassengerId %in% train_set$PassengerId))
glimpse(test_set)
```

##Exercise 3

<p><i>Our target is called Survived. First, fit a logistic regression model using Pclass, Sex, Fare as your three
features. Fit the model using the glm() function.</p></i>

```{r}

# Fit a model with intercept only
mod_1 <- glm(Survived ~ (Pclass+Sex+Fare), data = train_set, family = 'binomial')


# take a look at the features and coefficients
tidy(mod_1)

```

<i><p>After fitting the model, output the coefficients using the broom package and answer these questions:</p></i>

<p> <i>How would you interpret the coefficients?</i></p>
<p>For every additional ticket class increase (i.e. a lower class, since first class is the "best"), the odds of someone survivng decrease by 0.87.</p>
<p>If a passenger is male, the odds of them survivng decrease by 2.84.</p>

<p> <i>Are the features significant?</i></p>
All are statistically significant, looking at the p-value.

##Exercise 4


```{r}


library(rpart)
library(partykit)
tree_mod <- rpart(Survived ~ (Pclass+Sex+Fare), data = train_set)
plot(as.party(tree_mod))

```

<p>Describe in words one path a Titanic passenger might take down the tree. (Hint: look at your tree,
choose a path from the top to a terminal node, and describe the path like this - a male passenger who
paid a fare > 30 and was in first class has a high probability of survival) </p>

<p> Females with in a class less than 2.5 have a high probability of survival. </p>

<p>Does anything surprise you about the fitted tree?</p>

<p> It surprises me that fare only "matters" for females with a class greater or equal to 2.5. </p>


##Exercise 5



```{r}


test_logit <- predict(mod_1, newdata = test_set, type = 'response')
test_tree <- predict(tree_mod, newdata = test_set)[,2]


library(ROCR)
# create the prediction objects for both models
pred_logit <- prediction(predictions = test_logit, labels = test_set$Survived)
pred_tree <- prediction(predictions = test_tree, labels = test_set$Survived)
# get the FPR and TPR for the logistic model
# recall that the ROC curve plots the FPR on the x-axis
perf_logit <- performance(pred_logit, measure = 'tpr', x.measure = 'fpr')
perf_logit_tbl <- tibble(perf_logit@x.values[[1]], perf_logit@y.values[[1]])
# Change the names of the columns of the tibble
names(perf_logit_tbl) <- c('fpr', 'tpr')
# get the FPR and TPR for the tree model
perf_tree <- performance(pred_tree, measure = 'tpr', x.measure = 'fpr')
perf_tree_tbl <- tibble(perf_tree@x.values[[1]], perf_tree@y.values[[1]])
# Change the names of the columns of the tibble
names(perf_tree_tbl) <- c('fpr', 'tpr')
# Plotting function for plotting a nice ROC curve using ggplot
plot_roc <- function(perf_tbl) {
p <- ggplot(data = perf_tbl, aes(x = fpr, y = tpr)) +
geom_line(color = 'blue') +
geom_abline(intercept = 0, slope = 1, lty = 3) +
labs(x = 'False positive rate', y = 'True positive rate') +
theme_bw()
return(p)
}
# Create the ROC curves using the function we created above
plot_roc(perf_logit_tbl)
plot_roc(perf_tree_tbl)
```
























