---
title: "COMPSCIX 415.2 Homework 3"
author: "Shannon Hamilton"
date: "2/13/2018"
output: 
  html_document:
  self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
```
<h2> 3.8.1 Exercises </h2>


<i>1.What is the problem with this plot? How could you improve it?</i> <p></p>
There are overlapping values (overplotting), which makes it difficult to see where there are larger groupings of data points. Jittering can improve this plot.<p></p>


```{r}
data("mpg")

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy), position = "jitter")

```

<i>2. What parameters to geom_jitter() control the amount of jittering?</i> <p></p>
Width controls the amount of hotizontal jitter, height controls veritcal jitter.<p></p>

<i>3.Compare and contrast geom_jitter() with geom_count()</i> <p></p>
geom_jitter add slight random variation to points so that they don't overlap, while geom_count makes the size of points that overlap bigger depending on the count of the number of points that overlap. <p></p>

<i>4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.</i> <p></p>
The default position is dodge, which places overlapping objects directly beside one another. <p></p><p></p>

```{r}
data("mpg")

ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(class, hwy, color = drv))

```

<h2> 3.9.1 Exercises </h2>

<i>1.Question</i> <p></p>
Answer.<p></p>

<i>1.Question</i> <p></p>
Answer.<p></p>

<i>1.Question</i> <p></p>
Answer.<p></p>