---
title: "COMPSCIX 415.2 Homework 2"
author: "Shannon Hamilton"
date: "2/6/2018"
output: 
  html_document:
  self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
```

<h2> 3.2.4 Exercises </h2>


<i>1. Run ggplot(data = mpg). What do you see?</i> <p></p>
A blank plot.<p></p>

<i>2. How many rows are in mpg? How many columns?</i><p></p>
There are 234 rows of data (i.e. observations) and 11 columns (i.e. variables).<p></p>

<i>3. What does the drv variable describe? Read the help for? mpg to find out.</i><p></p>

drv describes the driving type of the car, which takes on three potential values: front-wheel drive, reel-wheel drive and four-wheel drive. <p></p>

<i>4. Make a scatterplot of hwy vs cyl.</i><p></p>
See code.
<p></p>

<i>5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?</i><p></p>
The scatterplot for these variables isn't valuable because both of these variables are categorical where the categories are nominal. "r" being higher than "f" and "4" doesn't hae any meaning. Similarly, "2seater" isn't "lower" or "less than" than "minivan".<p></p>

```{r}
data("mpg")

ggplot(data = mpg)
glimpse(mpg)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))
```

<h2> 3.3.1 Exercises </h2>

<i>1. What’s gone wrong with this code? Why are the points not blue?</i><p></p>

The points are not blue because the "color = blue" is within the aesthetic argument brackets. To make the points blue, the aes brackets must be closed and the "color" argument must be within the geom_point brackets, but after the aes brackets.  <p></p>

<i>2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?</i><p></p>

The categorical variables are: model name, transmission type, driving type, type of car.
The continuous variables are: displacement, year, number of cylinders, city mpg, highway mpg.
<p></p>

<i>3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?</i><p></p>

When you map color to a continous variable, it creates a gradient. When you map size to a continous variable, it creates a gradient of size. Continuos variables cannot be used with shape.  <p></p>

```{r}
data("mpg")

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl, color=displ))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl, size=displ))

```

<i>4. What happens if you map the same variable to multiple aesthetics?</i><p></p>

If you map the same variable to multiple aesthetics, the graph will build with both aesthetics. <p></p>

```{r}
data("mpg")

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl, color=displ,size=displ))

```

<i>5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point) </i><p></p>

Stroke controls the boarder width of select shapes (# 21-25). <p></p>

<i>6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?</i><p></p>

If used with a condition, the aesthetic will create two types (e.g. two sizes) that map to the condition being true and another for the condition being false. <p></p>

```{r}
data("mpg")

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl, color=displ < 5))


ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl, size=displ < 5))

```

<h2> 3.5.1 Exercises </h2>

<i>4. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?</i><p></p>

```{r}
data("mpg")

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)

```

It is easier to see the trends by car type with the facets over color-coded car type. However, it is harder to see overall trends and compare across car types. As the data set gets bigger, the facet becomes easier to read than color. <p></p>

<i>5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol argument?</i><p></p>

The nrow function determines the number of rows in the facet. The ncol determines the number of columnms in the facet.The facet_grid automatically creates the columns and rows based on the number of unique values per variable. <p></p>

<h2> 3.6.1 Exercises </h2>

<i>1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?</i><p></p>

Bar charts use bar geoms, line charts use line geoms, boxplots use boxplot geoms and area chats use area geoms. <p></p>

<i>2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.</i><p></p>

I expect a series of scatterplots (one per car class) in two rows with displacement and mpg as the axes. <p></p>

```{r}
data("mpg")

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)

```

<i>3. What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?</i><p></p>

It removes the legend from a chart.It was used previously so that there wasn't a legend only on a single plot in a series.  <p></p>

<i>4. What does the se argument to geom_smooth() do?</i><p></p>

se adds/removes error bands. <p></p>

<i>5. Will these two graphs look different? Why/why not?</i><p></p>

They will look the same as they are both based on the same information. This information can be defined at the global level so that it doesn't need to be repeated in the code. <p></p>

<h2> 3.7.1 Exercises </h2>

<i>2. What does geom_col() do? How is it different to geom_bar()?</i><p></p>

geom_bar uses a count as its default stat. geom_col uses stat_identity: it leaves the data as is, so creates the height based on a variable from the data set. <p></p>

<h2>Other Questions </h2>

<i>1. Look at the data graphics at the following link: What is a Data Scientist. Please briefly critique the designer’s choices. What works? What doesn’t work? What would you have done differently? </i><p></p>

What Works: <p></p>
- Series of bar charts to represent the different education levels <p></p>
- Insight per chart called out <p></p>
<p></p>

<p></p>What Doesn't Work: <p></p>
- Colors are too similar making it harder to read / see differences <p></p>
- The "best source of new data science talent" chart is confusing in that the bars are different widths, but that doesn't seem to relresent anything <p></p>
- Confusing that most of the charts have to do with data scientists vs. BI, but in one of the later charts, the comparison is "normal data science" vs. "big data science"
- Inconsistency of grey = data science, blue = BI 
<p></p>

<p></p>What I would have done differently: <p></p>
- No pie charts! <p></p>
- More distinct colors <p></p>
- Consistency of colors to represent data scientist vs  BI <p></p>

<p></p>
