---
title: "COMPSCIX 415.2 Homework 3"
author: "Shannon Hamilton"
date: "2/13/2018"
output: 
  html_document:
  self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
library(nycflights13)
```
<h2> 3.8.1 Exercises </h2>


<i>1.What is the problem with this plot? How could you improve it?</i> <p></p>
There are overlapping values (overplotting), which makes it difficult to see where there are larger groupings of data points. Jittering can improve this plot.<p></p>


```{r}
data("mpg")

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy), position = "jitter")

```

<i>2. What parameters to geom_jitter() control the amount of jittering?</i> <p></p>
Width controls the amount of hotizontal jitter, height controls veritcal jitter.<p></p>

<i>3.Compare and contrast geom_jitter() with geom_count()</i> <p></p>
geom_jitter add slight random variation to points so that they don't overlap, while geom_count makes the size of points that overlap bigger depending on the count of the number of points that overlap. <p></p>

<i>4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.</i> <p></p>
The default position is dodge, which places overlapping objects directly beside one another. <p></p><p></p>

```{r}
data("mpg")

ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(class, hwy, color = drv))

```

<h2> 3.9.1 Exercises </h2>

<i>2.What does labs() do? Read the documentation.</i> <p></p>

labs() allows you to modify the labels of your plot - the chart title or x/y axis titles. <p></p>

<i>4. What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?</i> <p></p>

coord_fixed creates a specified ratio between the physical representation of data units on the axes. This helps to decrease the likelihood of misunderstanding from streatched out axies. geom_abline() adds a reference line to the data. <p></p>

```{r}
data("mpg")

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_abline() +
  coord_fixed()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() 

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  coord_fixed()

```

<h2> 4.4 Exercises </h2>

<i>1. Why does this code not work?</i> <p></p>
It doesn't return an answer because the user misspelled the object name (it isn't an "i" in "variable") from what they defined when they tried to return it. <p></p>

<i>2.Tweak each of the following R commands so that they run correctly:</i> <p></p>
See below. <p></p>

```{r}
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
filter(mpg, cyl == "8")

filter(diamonds, carat > 3)


```


<h2> 5.2.4 Exercises </h2> <p></p>
<i>1.Find flights that: </i> <p></p>
Had an arrival delay of two or more hours <p></p>


```{r}

data("flights")
(twohour_flights <- filter(flights, arr_delay >= 120))

```

<p></p Flew to Houston (IAH or HOU)<p></p>

```{r}

data("flights")
(houston_flights <- filter(flights, dest %in% c("HOU", "IAH")))

```

<p></p Were operated by United, American, or Delta <p></p>

```{r}

data("flights")
(UAD_flights <- filter(flights, carrier %in% c("UA", "DL", "AA")))

```

<p></p Departed in summer (July, August, and September)<p></p>
```{r}

data("flights")
(UAD_flights <- filter(flights, month %in% c(7, 8, 9)))

```
<p></p Arrived more than two hours late, but didn’t leave late. <p></p>

```{r}

data("flights")
(UAD_flights <- filter(flights, arr_delay >= 120, dep_delay <=0))

```

<p></p Were delayed by at least an hour, but made up over 30 minutes in flight. (IAH or HOU)<p></p>

```{r}

data("flights")
(UAD_flights <- filter(flights, dep_delay >=60, (dep_delay - arr_delay) >= 30))

```

<p></p Departed between midnight and 6am (inclusive). <p></p


<i>1.Question</i> <p></p>
Answer.<p></p>

```{r}

data("flights")
twohour_flights <- filter(flights, arr_delay >= 120)

```

<i>1.Question</i> <p></p>
Answer.<p></p>