---
title: "COMPSCIX 415.2 Homework 4"
author: "Shannon Hamilton"
date: "2/20/2018"
output: 
  html_document:
  self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
library(nycflights13)
```
<h2> 5.6.7 Exercises</h2>


<i>2.Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).</i><p></p>

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% count(dest)

#alternative
not_cancelled %>% 
   group_by(dest) %>% 
  summarise(n = n())

not_cancelled %>% count(tailnum, wt = distance)

#alternative
not_cancelled %>% 
   group_by(tailnum) %>% 
  summarise(n = sum(distance))

```



<i>4. For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.</i><p></p>

```{r}
flights %>% 
  group_by(dest) %>% 
  summarise(delay = sum(dep_delay, na.rm = TRUE))

flights %>% 
  group_by(tailnum, dest) %>% 
  summarise(proportion = mean(dep_delay[!is.na(dep_delay)] > 0))

```



<i>5. Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag() explore how the delay of a flight is related to the delay of the immediately preceding flight.</i><p></p>


```{r}

flights %>%
  select(year, month, day, hour, dest, dep_delay) %>%
  group_by(dest) %>%
  mutate(lagdelay = lag(dep_delay)) %>%
  filter(!is.na(lagdelay)) %>%
  arrange(dest) %>%
  summarize(lag_correlation = cor(dep_delay, lagdelay, use = "complete.obs"))

```
<p>There does appear to be a positive correlation - at least in some locations (e.g. IAH, MDW, MCO) - between the lagged previous flight and the current flighr delay. </p><p></p>


<i>6.Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?</i><p></p>

```{r}
data(flights)

flights %>%
  group_by(dest) %>%
  mutate(mean_air = mean(air_time, na.rm = TRUE), fast = ((air_time - mean_air)/mean_air))%>%
  select(tailnum, air_time, mean_air, fast)%>%
  arrange(fast)



```

<p>Comparing to the mean flight time, N947UW, N666DN and N14568 tail numbers are far below the flight time average by location. Tail numbers N703TW, N178DN and N5DBAA had the longest delay in the air. </p><p></p>

<h2> 10.5 Exercises </h2>
<i>1.How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).</i><p></p>

<p>Tibbles only print out the first few rows of data instead of all of the data like in a data frame. Tibbles also indicate the type of date at the top of the columns. </p><p></p>


<i>2.Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?</i><p></p>

<p>$ returns a response for the data frame, while the tibble does not when an incomplete column name is provided. The tibble approach decreases the likelihood of a mis-type resulting in pulling the wrong data. For the [ to select data, the data frame returns a vector. If you were feeding this into another argument you may need to know the size of the vector, which could lead to errors in downstream code.  </p><p></p>


```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df_tibble <- as_tibble(df)
df_tibble$x

```
```{r}
df <- data.frame(abc = 1, xyz = "a")
df_tibble <- as_tibble(df)
df[, "xyz"]
df_tibble[, "xyz"]

```

```{r}
df <- data.frame(abc = 1, xyz = "a")
df_tibble <- as_tibble(df)
df[, c("abc", "xyz")]
df_tibble[, c("abc", "xyz")]

```

<i>3.If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble? </i><p></p>

<p>You can utilize square brackets to extract the name of a variable stored in an object of a tibble. </p><p></p>

<i>6.What option controls how many additional column names are printed at the footer of a tibble?</i><p></p>

<p>From Help, n_extra decides the extra columns for which to print abbreviated information . <p><p></p>

<h2> 12.3.3 Exercises - return to # 2 </h2>

<i>2. Why does this code fail?</i><p></p>

```{r}
table4a %>% 
  gather("1999", "2000", key = "year", value = "cases")
```

<p>The Column names needed to be in quotes. </p><p></p>

<i>3.Why does spreading this tibble fail? How could you add a new column to fix the problem?</i><p></p>

```{r}

people <- tribble(
  ~name,             ~key,    ~value, ~year,
  #-----------------|--------|------|------
  "Phillip Woods",   "age",       45, 1990,
  "Phillip Woods",   "height",   186, 1990,
  "Phillip Woods",   "age",       50, 2000,
  "Jessica Cordero", "age",       37, 1990,
  "Jessica Cordero", "height",   156, 1990
)

spread(people,key,value)

```

<p>Phillip has two rows with two different ages, which is causing the error. Perhaps this data is missing a year that would make it all the rows unique. </p><p></p>

<i>4.Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?</i><p></p>

```{r}

preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)

gather(preg,sex, count, male, female) 
  
```

<p>The variables are gender, pregnant and count. We need to gather. </p><p></p>

<h2> Section 12.4.3 Exercises </h2>

<i>1.What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.</i><p></p>

<p>If sep is a character vector, "extra" controls what happens when there are too many pieces. There are three valid options: warn, drop or merge. If sep is a character vector, fill" controls what happens when there are not enough pieces. There are three valid options: warn, right and left. </p><p></p>

```{r}

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "warn")

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra="merge")

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra="drop")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "left")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "right")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill="warn")
  
```

<i>2.Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?</i><p></p>

<p>"remove" if it is true removes input columsn from the output data frame.You may use it if you want to create a new variable based on an old variable and also keep the old variable.  </p><p></p>


<h2> Baby Names Exercises </h2>
```{r}
file_path <- '/Users/shamilto/Downloads/baby_names.txt'
baby <-read_delim(file_path,delim="|")
glimpse(baby)
```

```{r}
saveRDS(baby, file = "/Users/shamilto/Downloads/baby_names.rds")
```

```{r}
file_path2 <- '/Users/shamilto/Downloads/baby_names.rds'
baby2 <-read_rds(file_path2)
glimpse(baby2)
```