---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Shannon Hamilton"
date: "3/1/2018"
output: 
  html_document: 
    toc: yes
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

<i> Github Repository:https://github.com/shannonleehamilton/compscix-415-2-assignments.git </i><p></p>
  

##The tidyverse packages


<i>1. Can you name which package is associated with each task below?</i><p></p>

<ul>
<li>Plotting - ggplot2 </li>
<li>Data munging/wrangling - dplyr </li>
<li>Reshaping (speading and gathering) data -  tidyr </li>
<li>Importing/exporting data - readr </li>
</ul>

<i>2. Now can you name two functions that you’ve used from each package that you listed above for these tasks?</i><p></p>

<ul>
<li>Plotting - ggplot2: geom_point, geom_boxplot </li>
<li>Data munging/wrangling - dplyr: filter, select </li>
<li>Reshaping (speading and gathering) data -  tidyr: </li>
<li>Importing/exporting data - readr:read_delim, write_delim </li>
</ul>

##R Basics

<i>1. Fix this code with the fewest number of changes possible so it works:</i><p></p>

```{r}
My_data.name___is.too00ooLong <- c( 1 , 2 , 3 )

```
<p></p>
<i>2. Fix this code so it work:</i><p></p>

```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
```

<p></p>
<i>3.Look at the code below and comment on what happened to the values in the vector:</i><p></p>

```{r}
my_vector <- c(1, 2, '3', '4', 5)
my_vector
```

<p>Since some of the values in the vector are indicated to be a string (by putting them in quotations), R converts all of the values in the vector strings instead of recognizing them as numbers. </p>

##Data import/export

<i>1. Download the rail_trail.txt file from Canvas (in the Midterm Exam section here) and successfully import it into R. Prove that it was imported successfully by including your import code and taking a glimpse of the result</i><p></p>

```{r}
file_path <- '/Users/shamilto/Downloads/rail_trail.txt'
rail <-read_delim(file_path,delim="|")
glimpse(rail)
```

<i>2. Export the file into an R-specific format and name it “rail_trail.rds”. Make sure you define the path correctly so that you know where it gets saved. Then reload the file. Include your export and import code and take another glimpse.</i><p></p>

```{r}
saveRDS(rail, file = "/Users/shamilto/Downloads/rail_trail.rds")
```

```{r}
file_path2 <- '/Users/shamilto/Downloads/rail_trail.rds'
rail2 <-read_rds(file_path2)
glimpse(rail2)
```

##Visualization - Q2, Q3

<i>1.Critique this graphic: give only three examples of what is wrong with this graphic. Be concise.</i><p></p>

<p></p>

<ul>
<li>Comaring gender and age in the same axis is confusing and potentially misleading as there is potentially an interaction effect between gender and age in peoples' opinion on this topic </li>
<li>It is hard to compare the relative size of the bubbles and the colors aren't much help either!</li>
<li>It is confusing there isn't any deliniation between yes and no outside of the titles </li>
</ul>

<p></p>

<i>2. Reproduce this graphic using the diamonds data set.</i><p></p>

```{r}
 data("diamonds")

ggplot(diamonds) + 
  geom_boxplot(mapping = aes(x = cut, y = carat, fill=color)) +
  xlab("Cut of Diamond") +
  ylab("Carat of Diamond") +
  coord_flip()


```

<p></p>

<p></p>

<i>3. The previous graphic is not very useful. We can make it much more useful by changing one thing about it. Make the change and plot it again.</i><p></p>

<p>It is hard to map/ compare the "color" variable fill within the plot. Side-by-side boxplots to compare.</p>

```{r}
 data("diamonds")

ggplot(diamonds) + 
  geom_boxplot(mapping = aes(x = cut, y = carat, fill=color)) +
  xlab("Cut of Diamond") +
  ylab("Carat of Diamond") +
  coord_flip()


```


##Data munging and wrangling 

<i>1. Is this data “tidy”? If yes, leave it alone and go to the next problem. If no, make it tidy</i>

<p>This data is not tidy. "Cases" and "Population" should be their own columns.</p>

```{r}
  table2
  spread(table2, key = type, value = count)
```


<i>2.Create a new column in the diamonds data set called price_per_carat that shows the price of each
diamond per carat (hint: divide). Only show me the code, not the output.</i><p></p>

```{r}
  data("diamonds")

  diamonds %>%
  mutate(price_per_carat = (price/carat))  %>%
  arrange(-price_per_carat)
  
```

<i>3.For each cut of diamond in the diamonds data set, how many diamonds, and what proportion, have a price > 10000 and a carat < 1.5? There are several ways to get to an answer, but your solution must use the data wrangling verbs from the tidyverse in order to get credit.</i><p></p>

  
```{r}
  data("diamonds")

  (select_count <-  diamonds %>%
  group_by(cut) %>%
  filter(carat > 1 & price > 10000) %>%
  select(cut, price, carat) %>%
  summarise(select_count = n()))

(count <- diamonds %>%
  group_by(cut) %>%
  select(cut, price, carat) %>%
  summarise(count = n()))



```

<i>3.a) Do the results make sense? Why?</i>

<p>I expected that the proportion of diamonds that meet our critera would increase with the cut quality, since  I would have expected larger expensive diamonds to have a better cut. However, the proportion that meet our critera is around the same (8-9%) across all the cuts except for "premium"</p>

<i>3.b) Do we need to be wary of any of these numbers? Why?</i><p></p>

<p>I would be wary of the "premium" information given its percentage of our select diamonds is much higher than the other categories of cut (13%). I also am warry that this sample seems to have a lot of "ideal" and "premium" diamonds, while I would imagine that the higher the cut quality, the smaller the number of diamonds available (the opporiste trend of this data).</p>

<h2>EDA - Q5</h2>

<i> Take a look at the txhousing data set that is included with the ggplot2 package and answer these questions:</i><p></p>
mutate(proportion_missing= ((count(is.na(sales)))/count(sales)))


txhousing %>%
group_by(city) %>%
mutate(missing= count(is.na(sales)) %>%
select(city,sales)

```{r}
data("txhousing")
  
#Question 1
txhousing %>%
 count(year)

#Question 2

 txhousing %>%
  summarise(num_cities = n_distinct(city))

#Question 3
txhousing %>%
select(city,year,month,sales) %>%
 arrange(-sales)

#Question 4
txhousing %>%
summarize(listing_sales_corr = cor(listings, sales, use = "complete.obs"))

#Question 5


#INSERT CODE
       
#Question 6

filtered_housing <-txhousing %>%
group_by(city,year, month)%>%
filter(sales>500) 

ggplot(data = filtered_housing) + 
  geom_point(mapping = aes(x = month, y = median, color=city))+
  facet_wrap(~ city, nrow = 5)

```
```{r}
filtered_housing <-txhousing %>%
group_by(city,year, month)%>%
filter(sales>500) 

ggplot(data = filtered_housing) + 
  geom_point(mapping = aes(x = month, y = median, color=city))+
  facet_wrap(~ city, nrow = 5)

```

<i> 1. During what time period is this data from?</i>
<p> This data is from 2000 - 2015. </p><p></p>

<i> 2. How many cities are represented?</i>
<p>46 cities are represented.</p><p></p>

<i> 3. Which city, month and year had the highest number of sales?</i>
<p> Houston in July 2015 has the highest number of sales.</p><p></p>

<i> 4. What kind of relationship do you think exists between the number of listings and the number of sales?
Check your assumption and show your work.</i>
<p>My hypothesis is that there is a positive relationship between the number of listings and the number of sales. I checked my assumption by compusting the correlation between these two variables and was correct. The correlation is very high at 0.92. </p><p></p>

<i> 5. What proportion of sales is missing for each city?</i>
<p></p>

<i> 6. Looking at only the cities and months with greater than 500 sales: (a) Are the distributions of the median sales price (column name median), when grouped by city,different? The same? Show your work. (b)Any cities that stand out that you’d want to investigate further? (c) Why might we want to filter out all cities and months with sales less than 500?</i>

<p>The distribution of median sales is different by city in terms of the min/max of values, the variation of those values, and the number of median sales > 500. Corpus Christie has few values represented.Fort Bend has a larger variance in median prices compared to El Paso. Median prices in Collin Country are generally higher than Fort Worth We may want to filter out sales less than 500 to remove small outlier months that could skew our data interpretation. </p><p></p>







