---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Shannon Hamilton"
date: "3/1/2018"
output: 
    html_document:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
```

**ADD TABLE OF CONTENTS**
<h2> The tidyverse packages</h2>


<i>1. Can you name which package is associated with each task below?</i><p></p>

<ul>
<li>Plotting - ggplot2 </li>
<li>Data munging/wrangling - dplyr </li>
<li>Reshaping (speading and gathering) data -  tidyr </li>
<li>Importing/exporting data - readr </li>
</ul>

<i>2. Now can you name two functions that you’ve used from each package that you listed above for these tasks?</i><p></p>

<ul>
<li>Plotting - ggplot2: geom_point, geom_boxplot </li>
<li>Data munging/wrangling - dplyr: filter, select </li>
<li>Reshaping (speading and gathering) data -  tidyr: </li>
<li>Importing/exporting data - readr:read_delim, write_delim </li>
</ul>

<h2>R Basics</h2>

<i>1. Fix this code with the fewest number of changes possible so it works:</i><p></p>

```{r}
My_data.name___is.too00ooLong <- c( 1 , 2 , 3 )

```
<p></p>
<i>2. Fix this code so it work:</i><p></p>

```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
```

<p></p>
<i>3.Look at the code below and comment on what happened to the values in the vector:</i><p></p>

```{r}
my_vector <- c(1, 2, '3', '4', 5)
my_vector
```

<p>Since some of the values in the vector are indicated to be a string (by putting them in quotations), R converts all of the values in the vector strings instead of recognizing them as numbers. </p>

<h2>Data import/export</h2>

<i>1. Download the rail_trail.txt file from Canvas (in the Midterm Exam section here) and successfully import it into R. Prove that it was imported successfully by including your import code and taking a glimpse of the result</i><p></p>

<p></p>

<i>2. Export the file into an R-specific format and name it “rail_trail.rds”. Make sure you define the path correctly so that you know where it gets saved. Then reload the file. Include your export and import code and take another glimpse.</i><p></p>

<p></p>



